Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-17 tex/scam/summary.tex
a

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$.  Then $p$
  satisfies the \emph{boundary flatness condition} if for all $j$, and
  for all $\mathbf{x}_{-j}$,
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

@@

\begin{stheorem}
\label{thm:summary_acdc_population}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ that satisfies the boundary flatness
condition. Suppose that $f$ is convex and continuously twice-differentiable on an open set around $C$.
and that $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min_{\{f_k\},\mu} \Big \{
\E\Bigl( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \Bigr)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions, and, with respective to $f^*_k$'s from above, define
\begin{equation}
g^*_k = \arg\min_{g_k} \Big\{ \E\Bigl( f(X) - \mu^* - 
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k(X_k) \Bigr)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
functions.  Then $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{stheorem}

@@

Given samples $(y_i, X_i)$, AC/DC becomes the following optimization.
\begin{align*}
\{\hat{f}_k\}_{k=1}^p &= \arg\min_{\{f_k \in \mathcal{C}^1\}} \frac{1}{n}\sum_{i=1}^n \left(
        y_i - \bar{y} - \sum_{k=1}^p f_k(X_{ik}) \right)^2 + \lambda \sum_{k=1}^p \| f_k \|_\infty \\
\forall k, \, \hat{g}_k &= \arg\min_{g_k \in \mathcal{C}^1} \frac{1}{n} \sum_{i=1}^n \left(
    y_i - \bar{y} - \sum_{k' \neq k} \hat{f}_{k'}(X_{ik'}) - g_k(X_{ik}) \right)^2 + \lambda \| g_k \|_\infty
\end{align*}
where $\bar{y}$ is the empirical mean of $y$. Our estimate of the relevant variables is $\hat{S} = \{ k \,:\, \| \hat{f}_k \| > 0 \trm{ or } \| \hat{g}_k \| > 0\}$.

@@

Our main result is the following.
\begin{stheorem}
Suppose assumptions A1-A5 hold. Let $\{\hat{f}_i\}$ be any AC solution and
let $\{\hat{g}_k\}$ be any DC solution, both estimated with 
regularization parameter $\lambda$ scaling as
$\lambda = \Theta \Big( s\tilde{\sigma} \sqrt{\frac{1}{n} \log^2 np} \Big)$. 
Suppose in addition that
\begin{gather}
\alpha_f/\tilde{\sigma} \geq c B^2 \sqrt{\frac{s^5}{n^{4/5}} \log^2
  np}\\
\alpha_g^2/\tilde{\sigma} \geq c B^4 \sqrt{\frac{s^5}{n^{4/5}}
  \log^2 2np}.
\end{gather} 
where $\tilde{\sigma} \equiv \max(\sigma, B)$ and $c$ is a constant dependent only on $b, c_1$.

Then, for sufficiently large $n$, with probability at least $1-\frac{1}{n}$:
\begin{align*}
\hat{f}_k \neq 0 \trm{ or } \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and } \hat{g}_k = 0 & \trm{ for all } k \notin S.
\end{align*}

@@

We remark that \citet{dalalyan:12} show that, even with the product distribution,
 under traditional smoothness
constraints, variable selection is achievable only if $n > O(e^s)$. 
Here we demonstrate that convexity yields the scaling $n =
O(\textrm{poly}(s))$.

>>>>

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$. Then $p$
  satisfies the \emph{boundary flatness condition} if it satisfies
  certain regularity conditions (see the precise statement in
  Definition~\ref{defn:boundary-point}) and if for all $j$, and for
  all $\mathbf{x}_{-j}$,
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

@@

\begin{stheorem}
\label{thm:summary_acdc_population}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ that satisfies the boundary flatness
condition. Suppose that $f$ is convex and continuously twice-differentiable on an open set around $C$
and that the derivatives $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min_{\{f_k\},\mu} \Big \{
\E\Bigl( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \Bigr)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions.  Using the $f^*_k$s above, define
\begin{equation}
g^*_k = \arg\min_{g_k} \Big\{ \E\Bigl( f(X) - \mu^* - 
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k(X_k) \Bigr)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
functions.  Then $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{stheorem}

@@

Given samples $(y_i, X_i)$, AC/DC becomes the following optimization:
\begin{align*}
\{\hat{f}_k\}_{k=1}^p &= \arg\min_{\{f_k \in \mathcal{C}^1\}} \frac{1}{n}\sum_{i=1}^n \left(
        y_i - \bar{y} - \sum_{k=1}^p f_k(X_{ik}) \right)^2 + \lambda \sum_{k=1}^p \| f_k \|_\infty \\
\forall k, \, \hat{g}_k &= \arg\min_{g_k \in \mathcal{C}^1} \frac{1}{n} \sum_{i=1}^n \left(
    y_i - \bar{y} - \sum_{k' \neq k} \hat{f}_{k'}(X_{ik'}) - g_k(X_{ik}) \right)^2 + \lambda \| g_k \|_\infty
\end{align*}
where $\bar{y}$ is the empirical mean of $y$. Our estimate of the relevant variables is $\hat{S} = \{ k \,:\, \| \hat{f}_k \| > 0 \trm{ or } \| \hat{g}_k \| > 0\}$.

@@

Our main result is the following.
\begin{stheorem}
Suppose assumptions A1-A5 hold. Let $\{\hat{f}_i\}$ be any AC solution and
let $\{\hat{g}_k\}$ be any DC solution, both estimated with 
regularization parameter $\lambda$ scaling as
$\lambda = \Theta \Big( s\tilde{\sigma} \sqrt{\frac{1}{n} \log^2 np} \Big)$. 
Suppose in addition that
\begin{gather}
\alpha_f/\tilde{\sigma} \geq c B^2 \sqrt{\frac{s^5}{n^{4/5}} \log^2
  np}\\
\alpha_g^2/\tilde{\sigma} \geq c B^4 \sqrt{\frac{s^5}{n^{4/5}}
  \log^2 2np}.
\end{gather} 
where $\tilde{\sigma} \equiv \max(\sigma, B)$ and $c$ is a constant dependent only on $b, c_1$.
Then, for sufficiently large $n$, with probability at least $1-\frac{1}{n}$,
\begin{align*}
\hat{f}_k \neq 0 \trm{ or } \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and } \hat{g}_k = 0 & \trm{ for all } k \notin S.
\end{align*}

@@

We remark that \citet{dalalyan:12} show that, even under the product distribution,
variable selection is achievable under traditional smoothness constraints only if $n > O(e^s)$. 
Here we demonstrate that convexity yields the scaling $n =
O(\textrm{poly}(s))$.
